Notes on conformal manifolds

### 2025-05-21 Draft_MLFlatDirections.tex
Added new run.

% ---- COMMENTS ----
\newcommand{\ce}[1]{\marginpar{\parbox{\marginparwidth}{\boldmath $\Longleftarrow$}}{\boldmath\bfseries (ce: #1)}}
\newcommand{\gl}[1]{\marginpar{\parbox{\marginparwidth}{\boldmath $\Longleftarrow$}}{\boldmath\bfseries (gl: #1)}}
\newcommand{\bd}[1]{\marginpar{\parbox{\marginparwidth}{\boldmath $\Longleftarrow$}}{\boldmath\bfseries (bd: #1)}}

@@

		We then perform gradient descent on the points using TensorFlow’s automatic differentiation. The loss function is defined as:
		\begin{equation}
			\mathcal{L} = \sum_{i = 1}^{n_{points}} || \nabla V(\vec{x}^{,i})||^2
		\end{equation}
		where $\vec{x}^{\,i} = (x^i_1,x^i_2,x^i_4,x^i_8,x^i_{10})$ denotes the data points. To do the gradient descent, we have noted throught trial en error that reinitializing the optimizer after a periodic amount of steps were significantly speeding up the convergence. This is similar to what has been proposed in \cite{loshchilov2017sgdrstochasticgradientdescent} where they do warm restarts. However, we have noted that by doing so, the convergence was slower than just reinitializing the optimizer. We interpret this by arguying that when we reinitialize the optimizer, it forgets about its passed. Therefore, the effective learning rate used by the optimizer is truly the one passed, and not one adjusted by the optimizer. The loss function as well as the learning rate history are presented in Fig. \ref{label}
		%We use the Adam optimizer with a learning rate of $10^{-2}$ and run the gradient descent for $10,000$ epochs to ensure convergence.

%		\begin{figure}
%			\includegraphics[scale = 0.5]{imagefile}
%			\caption{Loss function for the gradient descent as well as the learning rate schedule, fixed with \texttt{tf.keras.optimizers.schedules.CosineDecayRestarts}, with parameters \texttt{initial_learning_rate = 0.01}, \texttt{first_decay_steps = 400}, \texttt{t_mul = 1.0} , \texttt{m_mul = 1.0}}
%		\end{figure}

@@

	  Mean loss with regularisation for particles reproducing a target: $285 \pm 108$

	  Mean loss without regularisation for particles reproducing a target: $33 \pm 24$
  		\item For 100 runs with parameters
		\begin{equation}
			\begin{gathered}
				\texttt{max degree} = 4, \quad \texttt{max num monomials} = 6, \quad n_{\rm iter} = 1000, \quad n_{\rm particles} = 1000, \\
				\texttt{sample size} = 10000, \quad {\rm reg} = 10^{3}, \quad \beta = 10^{-10} + \left(\frac{i}{n_{\rm iter}}\right)^{5}, \quad \sigma = 0.1, \\
				p_{\rm add} = 0, \quad p_{\rm remove} = 0, \quad p_{\rm modify} = 0.5, \quad p_{\rm multiply} = 0.25, \quad p_{\rm divide} = 0.25,
			\end{gathered}
		\end{equation}
		the polynomials are found at the following frequencies
	  \begin{equation}
	  	\begin{tabular}{l|cccccccc}
	  		& $p_{1}$ & $p_{2}$ & $p_{3}$ & $p_{4}$ & $p_{5}$ & $p_{6}$ & $p_{7}$ & $\varnothing$ \\\hline\hline
	  		Frequency & 69\% & 11\% & 7\% & 12\% & 0\% & 4\% & 0\% & 13\%
	  	\end{tabular}
	  \end{equation}
	  The dynamics of the appearance of the target polynomials is plotted in fig.~\ref{fig:sample10000beta5}.

@@

	  Mean loss with regularisation for particles reproducing a target: $282 \pm 91$

	  Mean loss without regularisation for particles reproducing a target: $35 \pm 24$

@@

	  Mean loss with regularisation for particles reproducing a target: $310 \pm 115$

	  Mean loss without regularisation for particles reproducing a target: $58 \pm 40$

	  \begin{figure}[h!]
	  	\centering
	  	\begin{tikzpicture}
	  		\draw (0,0) node (fig1) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_19_at_06_01_21.pdf}};
	  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
	  		\draw (0,-8) node (fig2) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_19_at_06_01_21_detailed.pdf}};
	  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
	  	\end{tikzpicture}
	  	\caption{Evolution of the mean numbers of particles reproducing a target polynomials and its $1\sigma$ deviation for the parameters~\eqref{eq:sample10000beta3}. (b) Evolution of the mean numbers of particles reproducing each targets polynomial. The vertical lines indicate when there are an average 5 particles reproducing a given polynomial, and the value of $\beta$ at that moment.}
	  	\label{fig:sample10000beta3}
	  \end{figure}

@@

	  \begin{figure}[h!]
	  	\centering
	  	\begin{tikzpicture}
	  		\draw (0,0) node (fig1) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_20_at_08_01_44.pdf}};
	  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
	  		\draw (0,-8) node (fig2) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_20_at_08_01_44_detailed.pdf}};
	  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
	  	\end{tikzpicture}
	  	\caption{Evolution of the mean numbers of particles reproducing a target polynomials and its $1\sigma$ deviation for the parameters~\eqref{eq:sample1000}. (b) Evolution of the mean numbers of particles reproducing each targets polynomial. The vertical lines indicate when there are an average 5 particles reproducing a given polynomial, and the value of $\beta$ at that moment.}
	  	\label{fig:sample1000}
	  \end{figure}
  \end{enumerate}

@@

%For every couples $(p_{i},p_{j})$ of polynomials in eq.~\eqref{eq:pols}, the system
%\begin{equation}
%	\begin{cases}
%		p_{i} = 0, \\
%		p_{j} = 0,
%	\end{cases}
%\end{equation}
%gives the same solution \ce{That's not true! For example, $x_{1}=x_{2}=0$ solves all polynomials but is not a sugra solution. Mention tests on the sugra?}

In the previous section, we introduced a numerical method that enabled symbolic regression, yielding a set of polynomials that vanish on our dataset, as presented in Eq.~\ref{eq:pols}. We can now solve the system
\begin{equation}\label{eq:solvepol}
	p_i = 0, \quad \forall i \in {1,\dots,7}
\end{equation}
which leads to the following expressions:
\begin{equation}
	\begin{cases}
		\displaystyle x_{1} = \frac{x_{2}}{\sqrt{2}}\,\frac{e^{x_{8}/2}}{e^{x_{8}}-1}\,\Big(-x_{5}\,e^{x_{8}/2} + \sqrt{2-4\,e^{x_{8}}+e^{2x_{8}}\big(2+x_{10}^{2}\big)}\Big),\\[8pt]
		\displaystyle x_{4} = \frac{e^{-x_{8}/2}}{\sqrt{2}}\,\sqrt{2-4\,e^{x_{8}}+e^{2x_{8}}\big(2+x_{10}^{2}\big)}.
	\end{cases}
\end{equation}

Alternatively, the system can be recast as:
\begin{equation}\label{eq:rulex8x10}
	\begin{cases}
		\displaystyle \tilde{x}{8} = \frac{x{1}^{2}+x_{2}^{2}}{x_{2}^{2} + \big(x_{1}-x_{2}x_{4}\big)^{2}},\\[10pt]
		\displaystyle x_{10} = \sqrt{2},x_{4},\frac{x_{2}^{2} - x_{1}^{2}+x_{1}x_{2}x_{4}}{x_{1}^{2}+x_{2}^{2}}.
	\end{cases}
\end{equation}

As anticipated, this defines a three-parameter solution corresponding to a three-dimensional slice of the half-maximal supergravity scalar potential.

Let us make a few remarks at this stage. First, it is possible to find solutions to Eq.~\ref{eq:solvepol} which, however, are not solutions of $\nabla V = 0$. For instance $x_1 = x_2 = 0$; such cases are excluded, as they do not satisfy the stationarity condition and therefore do not correspond to valid physical solutions.

Secondly, one might argue that only two of the seven polynomials are sufficient to fully characterize the solution. That is, choosing any pair $(i,j) \in {1, \dots, 7}$ may suffice to extract a complete description. In practice, this is not entirely accurate. While such a pair can yield partial constraints—for example, recovering Eq.\ref{eq:rulex8x10}—it may also produce alternative (and potentially less general) parameterizations. Upon inspection, all such partial rules are found to be consistent with, and included in, the most general expressions given in Eq.\ref{eq:rulex8x10}.

The solution preserves a ${\rm U}(1)\times{\rm U}(1)$ gauged symmetry.

>>>>

% ---- COMMENTS ----
\newcommand{\ce}[1]{\marginpar{\parbox{\marginparwidth}{\boldmath $\Longleftarrow$}}{\boldmath\bfseries (ce: #1)}}
\newcommand{\gl}[1]{\marginpar{\parbox{\marginparwidth}{\boldmath $\Longleftarrow$}}{\boldmath\bfseries (gl: #1)}}

@@

		We then perform gradient descent on the points using TensorFlow’s automatic differentiation. The loss function is defined as:
		\begin{equation}
			\mathcal{L} = \sum_{i = 1}^{n_{points}} || \nabla V(\vec{x}^{,i})||^2
		\end{equation}
		where $\vec{x}^{\,i} = (x^i_1,x^i_2,x^i_4,x^i_8,x^i_{10})$ denotes the data points. We use the Adam optimizer with a learning rate of $10^{-2}$ and run the gradient descent for $10,000$ epochs to ensure convergence.

		% \begin{figure}
			% \includegraphics[scale = 0.5]{imagefile} % add loss history
			% \end{figure}

@@

	  Mean loss with regularisation for particles reproducing a target: $597 \pm 264$

	  Mean loss without regularisation for particles reproducing a target: $345 \pm 246$
  		\item For 100 runs with parameters
		\begin{equation}
			\begin{gathered}
				\texttt{max degree} = 4, \quad \texttt{max num monomials} = 6, \quad n_{\rm iter} = 1000, \quad n_{\rm particles} = 1000, \\
				\texttt{sample size} = 10000, \quad {\rm reg} = 10^{3}, \quad \beta = 10^{-10} + \left(\frac{i}{n_{\rm iter}}\right)^{5}, \quad \sigma = 0.1, \\
				p_{\rm add} = 0, \quad p_{\rm remove} = 0, \quad p_{\rm modify} = 0.5, \quad p_{\rm multiply} = 0.25, \quad p_{\rm divide} = 0.25,
			\end{gathered}
		\end{equation}
		the polynomials are found at the following frequencies
	  \begin{equation}
	  	\begin{tabular}{l|cccccccc}
	  		& $p_{1}$ & $p_{2}$ & $p_{3}$ & $p_{4}$ & $p_{5}$ & $p_{6}$ & $p_{7}$ & $\varnothing$ \\\hline\hline
	  		Frequency & 69\% & 11\% & 7\% & 12\% & 0\% & 4\% & 0\% & 13\%
	  	\end{tabular}
	  \end{equation}
	  The dynamics of the appearance of the target polynomials is plotted in fig.~\ref{fig:sample10000beta5}.

@@

	  Mean loss with regularisation for particles reproducing a target: $606 \pm 257$

	  Mean loss without regularisation for particles reproducing a target: $359 \pm 244$

@@

	  Mean loss with regularisation for particles reproducing a target: $854 \pm 427$

	  Mean loss without regularisation for particles reproducing a target: $602 \pm 417$

	  \begin{figure}[h!]
	  	\centering
	  	\begin{tikzpicture}
	  		\draw (0,0) node (fig1) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_19_at_06_01_21.pdf}};
	  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
	  		\draw (0,-8) node (fig2) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_19_at_06_01_21_detailed.pdf}};
	  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
	  	\end{tikzpicture}
	  	\caption{(a) Evolution of the mean numbers of particles reproducing a target polynomials and its $1\sigma$ deviation for the parameters~\eqref{eq:sample10000beta3}. (b) Evolution of the mean numbers of particles reproducing each targets polynomial. The vertical lines indicate when there are an average 5 particles reproducing a given polynomial, and the value of $\beta$ at that moment.}
	  	\label{fig:sample10000beta3}
	  \end{figure}

	  \item For 100 runs with parameters
		\begin{equation} \label{eq:sample10000beta3sigmavar}
			\begin{gathered}
				\texttt{max degree} = 4, \quad \texttt{max num monomials} = 6, \quad n_{\rm iter} = 1000, \quad n_{\rm particles} = 1000, \\
				\texttt{sample size} = 10000, \quad {\rm reg} = 10^{3}, \quad \beta = 10^{-10} + \left(\frac{i}{n_{\rm iter}}\right)^{3}, \quad \sigma = 0.5 - \frac{0.45}{1+\exp\big(10-20 i/n_{\rm iter}\big)}, \\
				p_{\rm add} = 0, \quad p_{\rm remove} = 0, \quad p_{\rm modify} = 0.5, \quad p_{\rm multiply} = 0.25, \quad p_{\rm divide} = 0.25,
			\end{gathered}
		\end{equation}
		the polynomials are found at the following frequencies
	  \begin{equation}
	  	\begin{tabular}{l|cccccccc}
	  		& $p_{1}$ & $p_{2}$ & $p_{3}$ & $p_{4}$ & $p_{5}$ & $p_{6}$ & $p_{7}$ & $\varnothing$ \\\hline\hline
	  		Frequency & 68\% & 17\% & 6\% & 9\% & 2\% & 1\% & 1\% & 11\%
	  	\end{tabular}
	  \end{equation}
	  The dynamics of the appearance of the target polynomials is plotted in fig.~\ref{fig:sample10000beta3sigmavar}.

	  Averaged number of particles reproducing a target in a successful run: $966 \pm 115$

	  Mean loss with regularisation for particles reproducing a target: $837 \pm 438$

	  Mean loss without regularisation for particles reproducing a target: $609 \pm 430$

	  \begin{figure}[h!]
	  	\centering
	  	\begin{tikzpicture}
	  		\draw (0,0) node (fig1) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_21_at_11_52_46.pdf}};
	  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
	  		\draw (0,-8) node (fig2) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_21_at_11_52_46_detailed.pdf}};
	  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
	  	\end{tikzpicture}
	  	\caption{(a) Evolution of the mean numbers of particles reproducing a target polynomials and its $1\sigma$ deviation for the parameters~\eqref{eq:sample10000beta3sigmavar}. (b) Evolution of the mean numbers of particles reproducing each targets polynomial. The vertical lines indicate when there are an average 5 particles reproducing a given polynomial, and the value of $\beta$ at that moment.}
	  	\label{fig:sample10000beta3sigmavar}
	  \end{figure}

@@

	  \begin{figure}[h!]
	  	\centering
	  	\begin{tikzpicture}
	  		\draw (0,0) node (fig1) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_20_at_08_01_44.pdf}};
	  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
	  		\draw (0,-8) node (fig2) {\includegraphics[scale=0.75]{Figures/Count_beta_history_2025_05_20_at_08_01_44_detailed.pdf}};
	  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
	  	\end{tikzpicture}
	  	\caption{(a) Evolution of the mean numbers of particles reproducing a target polynomials and its $1\sigma$ deviation for the parameters~\eqref{eq:sample1000}. (b) Evolution of the mean numbers of particles reproducing each targets polynomial. The vertical lines indicate when there are an average 5 particles reproducing a given polynomial, and the value of $\beta$ at that moment.}
	  	\label{fig:sample1000}
	  \end{figure}
  \end{enumerate}

@@

For every couples $(p_{i},p_{j})$ of polynomials in eq.~\eqref{eq:pols}, the system
\begin{equation}
	\begin{cases}
		p_{i} = 0, \\
		p_{j} = 0,
	\end{cases}
\end{equation}
gives the same solution \ce{That's not true! For example, $x_{1}=x_{2}=0$ solves all polynomials but is not a sugra solution. Mention tests on the sugra?}
\begin{equation}
	\begin{cases}
		\displaystyle x_{1} = \frac{x_{2}}{\sqrt{2}}\,\frac{e^{x_{8}/2}}{e^{x_{8}}-1}\,\Big(-x_{5}\,e^{x_{8}/2} + \sqrt{2-4\,e^{x_{8}}+e^{2x_{8}}\big(2+x_{10}^{2}\big)}\Big),\\[8pt]
		\displaystyle x_{4} = \frac{e^{-x_{8}/2}}{\sqrt{2}}\,\sqrt{2-4\,e^{x_{8}}+e^{2x_{8}}\big(2+x_{10}^{2}\big)}.
	\end{cases}
\end{equation}
We checked that this defines a 3-parameter solution of three-dimensional half-maximal supergravity. Equivalently:
\begin{equation}
	\begin{cases}
		\displaystyle \tilde{x}_{8} = \frac{x_{1}^{2}+x_{2}^{2}}{x_{2}^{2} + \big(x_{1}-x_{2}x_{4}\big)^{2}},\\[10pt]
		\displaystyle x_{10} = \sqrt{2}\,x_{4}\,\frac{x_{2}^{2} - x_{1}^{2}+x_{1}x_{2}x_{4}}{x_{1}^{2}+x_{2}^{2}}.
	\end{cases}
\end{equation}
The solution preserves a ${\rm U}(1)\times{\rm U}(1)$ gauged symmetry.
